---
title: "Earth Analytics First Map"
author: "Brian Yandell"
date: "2024-09-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is translated from
[first-map.ipynb](https://github.com/byandell-Tribal/first-map-template/blob/main/first-map.ipynb).
I first used
<https://www.codeconvert.ai/python-to-r-converter>,
which only got partway there.
Then I used 
[MS CoPilot](https://copilot.microsoft.com/)
(via UW-Madison license).

The 
[OpenStreetMap (OSM)](https://www.openstreetmap.org/)
data is accessed via `osmdata`. 
Static maps are plotted using `ggplot2` with `sf` bridging
with `geom_sf` for polygon and point layers.
OSM base map layer is added with 
`annotation_map_tile` from `ggspatial` package.
Interactive maps can be produces with package `tmap`.

Interactive map

```{r eval=FALSE}
# Load necessary libraries
library(sf)
library(ggplot2)
library(osmdata)
library(ggspatial)
library(tmap)
```

```{r}
# Search for locations by name - this might take a moment
osm_gdf <- osmdata::opq("UW-Madison") |>
  osmdata::add_osm_feature(key = "amenity", value = "university") |>
  osmdata::osmdata_sf()
osm_gdf
```

Plot the outline of UW-Madison.
This does not automatically overlay on a map.
                          
```{r}
# Plot OSM boundary
osm_map <- ggplot2::ggplot() +
  # Base map from OpenStreetMap
  ggspatial::annotation_map_tile(type = "osm", zoom = 14,
                                 progress = "none") +
  # OSM points
  ggplot2::geom_sf(data = osm_gdf$osm_points, 
          fill = "transparent", 
          alpha = 0.2, 
          color = "red", 
          size = 1) +
  # OSM multipolygons (there are no polygons)
  ggplot2::geom_sf(data = osm_gdf$osm_multipolygons, 
          fill = "transparent", 
          alpha = 0.2, 
          color = "red", 
          size = 2) +
  ggplot2::ggtitle("UW-Madison") +
  ggplot2::theme_minimal() +
  ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5)) +
  ggplot2::coord_sf()

# Display the map
print(osm_map)
```

```{r eval=FALSE}
# Save the map as a file to put on the web (not quite right)
ggplot2::ggsave("osm.html", plot = osm_map, device = "html")
```

Interactive plot using 
[tmap](https://cran.r-project.org/web/packages/tmap/vignettes/tmap-getstarted.html).
Display seems to require an intermediate step of `sf::st_make_valid()` to render properly.

```{r}
intmd <- sf::st_make_valid(osm_gdf$osm_multipolygons)
```

```{r warning=FALSE}
tmap::tmap_mode("view")
```

```{r}
tmap::tm_shape(intmd) +
  tmap::tm_basemap("OpenStreetMap") +
  tmap::tm_borders(col = "red", lwd = 2)
```
